---
title: "ESG Mock"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    #vertical_layout: fill
    #vertical_layout: scroll
    social: menu
    source_code: embed
runtime: shiny
---

```{r setup, include=FALSE}
rm(list = ls())
### load prepare the data or source them from data dictioanry
load("DATA/usEcoData2Test.Rdata")
######source("CODE/0_collect_prepare_data.R")
######source("CODE/0_collect_prepare_CSV.R")
library(dplyr)
library(tidyr)
library(ggplot2)
library(flexdashboard)
source("CODE/fun_generic.R")
```

Univariate Time Series
=======================================================================

Inputs {.sidebar}
-----------------------------------------------------------------------

```{r}
selectInput("var.name.sel", label = h3("Economic varibale"), 
  choices = as.character(dict.data$nameR), 
  selected = as.character(dict.data$nameR[1]))
### ***2DO following start date and end date should be linked with the chosen variable
dateRangeInput("dates.train.sel", label = h5("Traning sample date range"),
      start = "1970-01-01", 
      end = "2010-12-01")
```
HERE we explain the basic ideas, etc etc.



```{r}
selectInput("delta.sel", label = h5(paste0("Type of difference")), 
  choices = list("as-is"="asis", "log difference"="diffLog", "% return"="diffRelat", "absolute difference"="diffAbs"), 
  selected = "diffLog")
```
as-is: $\Delta x_t = x_t$; 
log diff: $\Delta x_t = \log(x_t)-\log(x_{t-1})$;
% return: $\Delta x_t = x_t / x_{t_1} - 1$;
abs diff: $\Delta x_t = x_t - x_{t_1}$;



```{r}
sliderInput(inputId = "ar.lag.sel", label = "Autoregressive lag", min = 0, max = 10, value = 1)
sliderInput(inputId = "df.lag.sel", label = "Differencing lag", min = 0, max = 10, value = 0)
sliderInput(inputId = "ma.lag.sel", label = "Moving-Average lag", min = 0, max = 10, value = 0)
```


```{r}
### ***2DO how to interacte training sample date range
dateRangeInput("dates.test.sel", label = h5("Testing sample date range"),
      start = "2011-01-01", 
      end = as.character(Sys.Date()))
```

```{r}
#############################BEGIN: Prepare Reactive Data #############################################
namVarb <- reactive({
  as.character(input$var.name.sel)
})

inp.date.beg <- reactive({
  as.Date(input$dates.train.sel[1])
})

inp.date.end <- reactive({
  as.Date(input$dates.train.sel[2])
})

date.beg.test <- reactive({
  as.Date(input$dates.test.sel[1])
})

date.end.test <- reactive({
  as.Date(input$dates.test.sel[2])
})


df.full.sample <- reactive({
  Check.StopIf(sum(names(lst.eco)==namVarb())!=1, 
               "The selected variable cannot be found in the data list. Check dict.data!!!")
  eval(parse(text=paste0(
   "tmp <- lst.eco$", namVarb()
  )))
  Check.ExistVarInDF(tmp, c("date", "value"))
  tmp
})

df.full.sample.delta <- reactive({
  tmp <- df.full.sample()
  Check.Unique(tmp$date)
  Check.StopIf(sum(is.na(tmp$value))>0, "Currently, we cannot handle NA in the eco variable!")
  if (!identical(sort(tmp$date), tmp$date)){
    tmp <- tmp[match(sort(tmp$date), tmp$date), ]
  }
  if(input$delta.sel == "asis"){
    res <- tmp
  } else if (input$delta.sel == "diffLog"){
    res <- tmp[-1, ]
    res$value <- log(tmp[-1, 2]) - log(head(tmp[, 2], dim(tmp)[1]-1))
  } else if (input$delta.sel == "diffRelat"){
    res <- tmp[-1, ]
    res$value <- tmp[-1, 2] / head(tmp[, 2], dim(tmp)[1]-1) -1
  } else if (input$delta.sel == "diffAbs"){
    res <- tmp[-1, ]
    res$value <- tmp[-1, 2] - head(tmp[, 2], dim(tmp)[1]-1)
  } else {
    stop("The chosen delta is not defined. Extend code here!")
  }
  res
})
### select the training sample value
df.eco2Use <- reactive({
  tmp <- df.full.sample()
  filter(tmp, date >= max(inp.date.beg(), min(tmp$date)),
         date <= min(inp.date.end(), max(tmp$date)))
})
### select training sample DELTA values
df.eco2Use.delta <- reactive({
  tmp <- df.full.sample.delta()
  filter(tmp, date > min(df.eco2Use()[, 1]),
         date <= max(df.eco2Use()[, 1]))
})
### US Recession stress period for training sample
df.date.UsRecess <- reactive({
  filter(dateUsRecess, monBeg >= min(df.eco2Use()[, 1]), monEnd <= max(df.eco2Use()[, 1]))
})
### select the testing sample value
df.test.sample <- reactive({
  tmp <- df.full.sample()
  filter(tmp, date >= max(date.beg.test(), min(tmp$date)),
         date <= min(date.end.test(), max(tmp$date)))
})
### select the testing sample delta value
df.test.sample.delta <- reactive({
  tmp <- df.full.sample.delta()
  filter(tmp, date > min(df.test.sample()[, 1]),
         date <= max(df.test.sample()[, 1]))
})
### US Recession stress period for testing sample
df.date.UsRecess.test <- reactive({
  filter(dateUsRecess, monBeg >= min(df.test.sample()[, 1]), monEnd <= max(df.test.sample()[, 1]))
})
########################################################################END: reactive data##


############### ANALYTIC #################
### 1. Stationarity test for the chosen variable
## ADF test 
res.adf.txt <- reactive({
  as.character(tseries::adf.test(as.numeric(df.eco2Use()$value), "stationary"))
})
## Phillips and Perron test
res.pp.txt <- reactive({
  as.character(PP.test(as.numeric(df.eco2Use()$value)))
})
## KPSS test
res.kpLevel.txt <- reactive({
  tseries::kpss.test(as.numeric(df.eco2Use()$value), null="Level")
})
res.kpTrend.txt <- reactive({
  tseries::kpss.test(as.numeric(df.eco2Use()$value), null="Trend")
})
### 2. Stationarity test for DELTAs
## ADF
res.adf.txt.delta <- reactive({
  as.character(tseries::adf.test(as.numeric(df.eco2Use.delta()$value), "stationary"))
})
## Phillips and Perron test
res.pp.txt.delta <- reactive({
  as.character(PP.test(as.numeric(df.eco2Use.delta()$value)))
})
## KPSS test
res.kpLevel.txt.delta <- reactive({
  tseries::kpss.test(as.numeric(df.eco2Use.delta()$value), null="Level")
})
res.kpTrend.txt.delta <- reactive({
  tseries::kpss.test(as.numeric(df.eco2Use.delta()$value), null="Trend")
})

mod.arima <- reactive({
  ar <- input$ar.lag.sel
  df <- input$df.lag.sel 
  ma <- input$ma.lag.sel
  ### quality checks
  Check.StopIf(max(c(ar, df, ma)) > (dim(df.eco2Use.delta())[1]+1), 
               "Chosen lag should be smaller than time series length!")
  arima(df.eco2Use.delta()$value, order = c(ar, df, ma))
})

df.arima.predict <- reactive({
  tmp.val <- df.test.sample()
  tmp.val$value <- as.numeric(tmp.val$value)
  tmp.delta <- df.test.sample.delta()
  tmp.delta.pred <- predict(mod.arima(), n.ahead=dim(tmp.delta)[1])
  df.pred.delta <- data.frame(date=tmp.delta$date, value=tmp.delta$value,
                              pred=as.numeric(tmp.delta.pred$pred), se=as.numeric(tmp.delta.pred$se))
  ### quality checks
  mapDe2Val <- match(tmp.val$date[-1], df.pred.delta$date)
  Check.StopIf(sum(is.na(mapDe2Val))>0, "Expect that delta dates should be all found")
  Check.Unique(mapDe2Val)
  df.pred.val <- tmp.val
  if(input$delta.sel == "asis"){ ##### 2DO: bug here! as-is does not work
    df.pred.val$pred <- c(df.pred.val$value[1], df.pred.delta$pred[mapDe2Val])
    df.pred.val$se <- c(df.pred.val$value[1], df.pred.delta$se[mapDe2Val])
  } else if (input$delta.sel == "diffLog"){
    df.pred.val$pred <- exp(cumsum(c(log(df.pred.val$value[1]), df.pred.delta$pred[mapDe2Val])))
    df.pred.val$se <- exp(cumsum(c(log(df.pred.val$value[1]), df.pred.delta$se[mapDe2Val])))
  } else if (input$delta.sel == "diffRelat"){
    df.pred.val$pred <- df.pred.val$value[1]*cumprod(1+c(0, df.pred.delta$pred[mapDe2Val]))
    df.pred.val$se <- df.pred.val$value[1]*cumprod(1+c(0, df.pred.delta$se[mapDe2Val]))
  } else if (input$delta.sel == "diffAbs"){
    df.pred.val$pred <- cumsum(c(df.pred.val$value[1], df.pred.delta$pred[mapDe2Val]))
    df.pred.val$se <- cumsum(c(df.pred.val$value[1], df.pred.delta$se[mapDe2Val]))
  } else {
    stop("The chosen delta is not defined. Extend code here!")
  }
  list(df.pred.delta=df.pred.delta, df.pred.val=df.pred.val)
})

### Residual stationary test results
## ADF test 
resid.adf.txt <- reactive({
  as.character(tseries::adf.test(as.numeric(mod.arima()$residuals), "stationary"))
})
## Phillips and Perron test
resid.pp.txt <- reactive({
  as.character(PP.test(as.numeric(mod.arima()$residuals)))
})
## KPSS test
resid.kpLevel.txt <- reactive({
  tseries::kpss.test(as.numeric(mod.arima()$residuals), null="Level")
})
resid.kpTrend.txt <- reactive({
  tseries::kpss.test(as.numeric(mod.arima()$residuals), null="Trend")
})
#########################################
```



Row {.tabset .tabset-fade}
-----------------------------------------------------------------------
### Plot of chosen variable
```{r}
#########PLOT over time
renderPlot({
    ggplot(df.eco2Use()) + 
    geom_line(aes(x=date, y=value), col="blue", size=2) + theme_bw()  + 
    geom_rect(data=df.date.UsRecess(), 
              aes(xmin=monBeg, xmax=monEnd, ymin=-Inf, ymax=+Inf), fill='red', alpha=0.2)
})
```

### Plot $\Delta$ 
```{r}
renderPlot({
  #par(mfrow=c(1,2))
  ### left plot
  ggplot(df.eco2Use.delta()) +
  geom_line(aes(x=date, y=value), col="blue", size=1) + theme_bw() +
  geom_rect(data=df.date.UsRecess(),
            aes(xmin=monBeg, xmax=monEnd, ymin=-Inf, ymax=+Inf), fill='red', alpha=0.2)
  ### right plot
  # ggplot(df.eco2Use.delta(), aes(x=value)) + 
  #   geom_histogram()
  #par(mfrow=c(1,1))
})
```


### Histogram $\Delta$
```{r}
renderPlot({
  ggplot(df.eco2Use.delta(), aes(x=value)) +
    geom_histogram(aes(y=..density..), fill="green", alpha=0.5) +
    geom_density(col="blue", size=1) +
    stat_function(fun=dnorm, args=list(mean=mean(df.eco2Use.delta()$value), sd=sd(df.eco2Use.delta()$value)), 
                  col="red", size=1) +
    scale_colour_manual(name = "", values = c("blue", "red"), 
                               breaks = c("bla", "blabla", "yada"), 
                               labels = c("kernel_est", "empirical density", "normal density")) +
    theme(legend.position="bottom") 
    #scale_fill_discrete(name="Type: ", labels=c("delta values", "Normal density")) + 
    #theme(legend.position="bottom") 
    ### ***2DO: fix the legend here!!!
})
```



### ACF $\Delta$
[**ACF**: correlation between values of the chosen variable at different time](https://en.wikipedia.org/wiki/Autocorrelation)

```{r}
### get ACF plot
renderPlot({
  acf.res <- acf(df.eco2Use.delta()$value, plot = F)
  plot(acf.res, main = "")
})
```

### PCF $\Delta$
[**PACF**: partial correlation between values of the chosen variable at different time](https://en.wikipedia.org/wiki/Partial_autocorrelation_function)

```{r}
### get ACF plot
renderPlot({
  pacf.res <- pacf(df.eco2Use.delta()$value, plot = F)
  plot(pacf.res, main = "")
})
```


### Stationarity 

#### Stationarity test on the chosen variable and its **$\Delta$**

- [ADF](https://en.wikipedia.org/wiki/Augmented_Dickey%E2%80%93Fuller_test) test ($H_0$ of unit root presence, i.e.,  $I(1)$ process vs. $H_A$ stationary)

- [PP](https://en.wikipedia.org/wiki/Augmented_Dickey%E2%80%93Fuller_test) test ($H_0$ of unit root presence, i.e.,  $I(1)$ process v.s., $H_A$ stationary)

- [KPSS](https://en.wikipedia.org/wiki/KPSS_test) test ($H_0$ of stationary or $I(0)$ process v.s. $H_A$ non-stationary): 
This test can be performed assuming the deterministic component of the variable is a constant or a constant with a linear trend.

```{r}
DT::renderDataTable({
  tab2Show <- data.frame(pVal.var=as.numeric(c(res.adf.txt()[4], res.pp.txt()[3],
                                         res.kpLevel.txt()[3], res.kpTrend.txt()[3])),
                         pVal.delta=as.numeric(c(res.adf.txt.delta()[4], res.pp.txt.delta()[3],
                                         res.kpLevel.txt.delta()[3], res.kpTrend.txt.delta()[3])),
                         stat.var = as.numeric(c(res.adf.txt()[1], res.pp.txt()[1],
                                         res.kpLevel.txt()[1], res.kpTrend.txt()[1])),
                         stat.delta=as.numeric(c(res.adf.txt.delta()[1], res.pp.txt.delta()[1],
                                         res.kpLevel.txt.delta()[1], res.kpTrend.txt.delta()[1])),
                         stringsAsFactors=FALSE)
  tab2Show <- tab2Show %>%
    mutate(test = c("ADF test", "PP test", "KPSS: constant", "KPSS: trend"),
           conclusion.var=ifelse(grepl("KPSS", test),
                             ifelse(pVal.var >= 0.05, "Stationary", "Not stationary"),
                             ifelse(pVal.var < 0.05, "Stationary", "Not stationary")),
           conclusion.delta=ifelse(grepl("KPSS", test),
                             ifelse(pVal.delta >= 0.05, "Stationary", "Not stationary"),
                             ifelse(pVal.delta < 0.05, "Stationary", "Not stationary")),
           pVal.var=round(pVal.var, 3), pVal.delta=round(pVal.delta, 3))
  tab2Show <- select(tab2Show, test, pVal.var, conclusion.var,  pVal.delta, conclusion.delta)
  names(tab2Show) <- c("Stationary tests", "varible: p-value", "variable: Conclusion",
                       "delta: p-value", "delta: Conclusion")
  tab2Show
},  options=list(dom = 't'), rownames = FALSE)
```


Row {.tabset .tabset-fade}
-----------------------------------------------------------------------
### ARIMA Fit on $\Delta$
- The ARIMA fit results are:
`r renderUI(HTML(gsub(" ", " ", paste(capture.output(mod.arima())[-c(1:4)], collapse = " <br/>"))))`

- Historical fit of the above model is:
```{r}
renderPlot({
    # fit.pred.res <- mod.arima()$residuals + as.numeric(df.eco2Use.delta()$value)
    # plot(df.eco2Use.delta()$date, df.eco2Use.delta()$value, col = "blue")
    # lines(df.eco2Use.delta()$date, fit.pred.res, col = "red", 
    #       ylim=c(min(fit.pred.res), max(fit.pred.res)))
  tmp <- data.frame(date=df.eco2Use.delta()$date, 
                    value=mod.arima()$residuals + as.numeric(df.eco2Use.delta()$value))
  ggplot(data=df.eco2Use.delta(), aes(x=date, y=value)) + 
    geom_line(col="blue", size=1, alpha=0.9) +
    geom_line(data=tmp, aes(x=date, y=value), col="red")
  })
```

- Diagnostic Overview 
```{r}
### get ACF plot
renderPlot({
  tsdiag(mod.arima())
}) #, height = 400, width = 600 
```

- Residual PACF plot
```{r}
### get PACF plot
renderPlot({
  pacf.res <- pacf(mod.arima()$residuals, plot = F)
  plot(pacf.res, main = "PACF plot of fitted model residuals")
})
```

### Residual plots 
Residual plot (over time) of the above model is:
```{r}
renderPlot({
  tmp <- data.frame(date=df.eco2Use.delta()$date, residuals=as.numeric(mod.arima()$residuals))
  ggplot(tmp, aes(x=date, y=residuals)) + 
    geom_point(col="blue", size=2) +
    geom_hline(yintercept = 0, col="red")
})
```


### Residual Normality
- QQ plot of residuals 
```{r}
### ***2D: add confidence iternvals
renderPlot({
  tmp <- data.frame(residuals=as.numeric(mod.arima()$residuals))
  ggplot(data=as.data.frame(qqnorm(tmp$residuals, plot=FALSE)), aes(x=x, y=y)) + 
    geom_point(size=2, col="blue") + 
    geom_smooth(method="lm", se=FALSE, col="red") + 
    scale_x_continuous(name="Theoretical quantiles") + 
    scale_y_continuous(name="Sample quantiles")
  # ggplot(tmp, aes(sample = residuals)) + 
  #   stat_qq(shape=1) 
  })
```


### Residual test
#### Stationary tests on residuals
```{r}
DT::renderDataTable({
  tab2Show <- data.frame(pVal.var=as.numeric(c(resid.adf.txt()[4], resid.pp.txt()[3], 
                                         resid.kpLevel.txt()[3], resid.kpTrend.txt()[3])),
                         stringsAsFactors=FALSE)
  tab2Show <- tab2Show %>%
    mutate(test = c("ADF test", "PP test", "KPSS: constant", "KPSS: trend"),
           conclusion.var=ifelse(grepl("KPSS", test), 
                             ifelse(pVal.var >= 0.05, "Stationary", "Not stationary"), 
                             ifelse(pVal.var < 0.05, "Stationary", "Not stationary")))
  tab2Show[, grepl("pVal", names(tab2Show))] <- round(tab2Show[, grepl("pVal", names(tab2Show))], 3)
  tab2Show <- select(tab2Show, test, pVal.var, conclusion.var)
  names(tab2Show) <- c("Stationary tests", "p-value", "Conclusion")
  tab2Show
},  options=list(dom = 't'), rownames = FALSE)
```

#### [Ljung-Box](http://en.wikipedia.org/wiki/Ljung%E2%80%93Box_test) test on residual randomness: 

- **Test hypothesis**: $H_0$ that residulas are not auto-correlated (i.e. the correlations in the population from which the sample is taken are 0, so that any observed correlations in the data result from randomness of the sampling process), v.s., $H_A$ that the residulas process statistically significant autocorrelation.

- **Statistics**: Ljung-Box test statistic of the residuals is `r renderUI(HTML(paste0(round(as.matrix(Box.test(mod.arima()$residuals, type="Ljung-Box")$statistic), 3), "")))` and the associated p-value is `r renderUI(HTML(paste0(round(as.matrix(Box.test(mod.arima()$residuals, type = "Ljung-Box")$p.value), 2), "")))`

```{r}
renderUI({
  res <- as.numeric(as.matrix(Box.test(mod.arima()$residuals, type = "Ljung-Box")$p.value))
  if(res >= 0.05){
    txt <- "Hence, the residuals of the fitted model exhibits statistically significant randomness"
  } else {
    txt <- "Hence, the residuals of the fitted model are NOT statistically-significant independently distributed."
  }
  HTML(paste0(txt, ""))
})
```


### Forecast $\Delta$
```{r}
#### Figure of forecast delta
renderPlot({
  df2Plot <- df.arima.predict()$df.pred.delta
  gObj <- ggplot(df2Plot) +
    geom_point(aes(x=date, y=value), col="blue", size=3, alpha=0.8)  +
    geom_line(data=df2Plot, aes(x=date, y=pred), col="black") +
    geom_point(data=df2Plot, aes(x=date, y=pred), col="black", size=3, alpha=0.5) +
    geom_line(data=df2Plot, aes(x=date, y=pred+2*se), col="red", size=1) +
    geom_line(data=df2Plot, aes(x=date, y=pred-2*se), col="red", size=1)
  if (nrow(df.date.UsRecess.test())==0) {
    gObj
  } else{
    gObj + geom_rect(data=df.date.UsRecess.test(),
            aes(xmin=monBeg, xmax=monEnd, ymin=-Inf, ymax=+Inf), fill='red', alpha=0.2)
  }
})
```


### Forecast Variable
```{r}
#### Figure of variable value 
renderPlot({
  df2Plot <- df.arima.predict()$df.pred.val
  gObj <- ggplot(df2Plot) + 
    geom_point(aes(x=date, y=value), col="blue", size=3, alpha=0.8)  +
    geom_line(data=df2Plot, aes(x=date, y=pred), col="black") +
    geom_point(data=df2Plot, aes(x=date, y=pred), col="black", size=3, alpha=0.5) 
  if (nrow(df.date.UsRecess.test())==0) {
    gObj
  } else{
    gObj + geom_rect(data=df.date.UsRecess.test(),
            aes(xmin=monBeg, xmax=monEnd, ymin=-Inf, ymax=+Inf), fill='red', alpha=0.2)
  }
})
```

Multivariate Time Series
=======================================================================

Inputs {.sidebar}
-----------------------------------------------------------------------