---
title: "ESG Mock"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    # vertical_layout: fill
    # vertical_layout: scroll
    # social: menu
    # source_code: embed
runtime: shiny
---

```{r setup, include=FALSE}
rm(list = ls())
### load prepare the data or source them from data dictioanry
load("DATA/usEcoData2Test.Rdata")
######source("CODE/0_collect_prepare_data.R")
######source("CODE/0_collect_prepare_CSV.R")
library(dplyr)
library(tidyr)
library(vars)
library(ggplot2)
library(plotly)
library(flexdashboard)
source("CODE/fun_generic.R")
### numeric input with model output
renderNumInputs <- function(x, inputId, label, nameDependVar) {
  Check.IsScalar(nameDependVar)
  Check.StopIf(!(is.vector(x) & is.vector(inputId) & is.vector(label)), "Must be all vector")
  Check.StopIf(!(length(x) == length(inputId) & length(label) == length(x)), "Must be same length")
  txtNumInput <- rep(NA, length(x))
  for (idx in 1:length(x)){
    txtNumInput[idx] <- paste0("numericInput(inputId = '", inputId[idx], "', label='", label[idx], "', value=", x[idx], ")")
  }
  txtNumInput <- paste0("h3(HTML(nameDependVar)), ", paste(txtNumInput, collapse = ","))
    
  eval(parse(text=paste0(
    "wellPanel(fluidRow(column(12,",
      txtNumInput,
    ")))"
  )))
}
```


Multivariate Time Series
=======================================================================

Inputs {.sidebar}
-----------------------------------------------------------------------

```{r}
checkboxGroupInput('sel.var.VAR.mod', label = h5("Variable chosen to VAR"),
                           names(lst.eco), selected = names(lst.eco)[grepl("GDP", names(lst.eco))])
### ***2DO following start date and end date should be linked with the chosen variable
dateRangeInput("sel.time.train", label = h5("Traning sample date range"),
      start = "1995-01-01", end = "2013-12-31")
### type of difference 
selectInput("delta.sel.VAR", label = h5(paste0("Type of difference / delta")),
  choices = list("as-is"="asis", "abs diff rates; log diff rest"="diffLogAbs", 
                 "log difference"="diffLog", "absolute difference"="diffAbs"), #"% return"="diffRelat", 
  selected = "diffLogAbs")
### type of value to VAR(p)
selectInput("data2Fit.sel.VAR", label = paste0("Type of value for VAR(p) model fit"),
  choices = list("Original values"="valOriginal"), # "Adjusted values based chosen delta (log value if log diff)"="valAjusted", "log of original values"="valLogOrg"), 
  selected = "valOriginal")
### lag for VAR(p) model fitting
sliderInput(inputId = "sel.VAR.lag", label = "p parameter: lag of VAR(p)", min = 0, max = 10, value = 1)
### ***2DO how to interacte training sample date range
dateRangeInput("sel.time.test", label = h5("Testing sample date range"),
      start = "2013-12-31", end = as.character(Sys.Date()))
```



```{r}
#############################BEGIN: Prepare Reactive Data #############################################
## select economic variables
namVarb.VAR <- reactive({
  as.character(input$sel.var.VAR.mod)
})
## date range for training sample
date.range.train <- reactive({
  data.frame(dateBeg=as.Date(input$sel.time.train[1]), dateEnd=as.Date(input$sel.time.train[2]), 
             stringsAsFactors=FALSE)
})
## date range for testing sample
date.range.test <- reactive({
  data.frame(dateBeg=as.Date(input$sel.time.test[1]), dateEnd=as.Date(input$sel.time.test[2]), 
             stringsAsFactors=FALSE)
})
## full sample according to the selected economic variables
df.full.sample.VAR <- reactive({
  Check.Unique(namVarb.VAR())
  Check.StopIf(length(setdiff(namVarb.VAR(), names(lst.eco)))!=0, 
               "The selected variables cannot be found in the data list. Check dict.data!!!")
  if(length(namVarb.VAR()) > 1){
    pos.eco <- which(names(lst.eco) %in% namVarb.VAR())
    ##### 1st: find the common dates across chosen economic variables!
    date.comm <- as.Date(intersect(lst.eco[[pos.eco[1]]]$date, lst.eco[[pos.eco[2]]]$date), 
                         origin="1970-01-01")
    if (length(pos.eco) > 2){
      for (idx in 3:length(pos.eco)){
        date.comm <- as.Date(intersect(date.comm, lst.eco[[pos.eco[idx]]]$date), origin="1970-01-01")
      }
    }
    tmp <- data.frame(date=date.comm, stringsAsFactors=FALSE)
    for (idx in pos.eco){
      eval(parse(text=paste0(
      "tmp$", names(lst.eco)[idx], " <- as.numeric(lst.eco[[", idx, 
          "]]$value[match(date.comm, lst.eco[[", idx,"]]$date)])"
      )))
    }
    tmp
  } else{
    stop("Must choose at least two economic variables!!!")
  }
})
## delta's of the full sample
df.full.sample.delta.VAR <- reactive({
  tmp <- df.full.sample.VAR()
  Check.Unique(tmp$date)
  Check.StopIf(sum(is.na(tmp$value))>0, "Currently, we cannot handle NA in the eco variable!")
  if (!identical(sort(tmp$date), tmp$date)){
    tmp <- tmp[match(sort(tmp$date), tmp$date), ]
  }
  posNonDate <- which(names(tmp) != "date")
  ### proced to calculate deltas
  if(input$delta.sel.VAR == "asis"){
    res <- tmp
  } else if (input$delta.sel.VAR == "diffLog"){
    res <- tmp[-1, ]
    res[, posNonDate] <- log(tmp[-1, posNonDate]) - log(head(tmp[, posNonDate], dim(tmp)[1]-1))
  } else if (input$delta.sel.VAR == "diffRelat"){
    res <- tmp[-1, ]
    res[, posNonDate] <- tmp[-1, posNonDate] / head(tmp[, posNonDate], dim(tmp)[1]-1) -1
  } else if (input$delta.sel.VAR == "diffAbs"){
    res <- tmp[-1, ]
    res[, posNonDate] <- tmp[-1, posNonDate] - head(tmp[, posNonDate], dim(tmp)[1]-1)
  } else if (input$delta.sel.VAR == "diffLogAbs") {
    res <- tmp[-1, ]
    ### 2DO: QUICK-FIX: we take absoluteDiff for variable name contains "rate"
    ### we take logDiff for variable for variable name does not contain "rate"
    ixRate <- grepl("rate.", names(tmp), ignore.case=T)
    ixNonRest <- (!grepl("rate.", names(tmp), ignore.case=T)) & (names(tmp) != "date")
    if (sum(ixRate) > 0) {
      res[, ixRate] <- tmp[-1, ixRate] - head(tmp[, ixRate], dim(tmp)[1]-1)
    }
    if (sum(ixNonRest) > 0) {
      res[, ixNonRest] <- log(tmp[-1, ixNonRest]) - log(head(tmp[, ixNonRest], dim(tmp)[1]-1))
    }
  } else {
    stop("The chosen delta is not defined. Extend code here!")
  }
  res
})
## traning sample according to the selected economic variables
df.train.sample.VAR <- reactive({
  tmp <- df.full.sample.VAR()
  Check.ExistVarInDF(tmp, "date")
  filter(tmp, date >= max(date.range.train()$dateBeg, min(tmp$date)),
         date <= min(date.range.train()$dateEnd, max(tmp$date)))
})
### traning sample aligned value (for non-rate variables, set all value starting at 100!)
df.train.sample.alignedVal.VAR <- reactive({
  tmp <- df.train.sample.VAR()
  res <- tmp
  ixNonRest <- (!grepl("rate.", names(tmp), ignore.case=T)) & (names(tmp) != "date")
  if (sum(ixNonRest) > 0){
    res[1, ixNonRest] <- 100
    tmpRet <- tmp[-1, ixNonRest] / head(tmp[, ixNonRest], dim(tmp)[1]-1) -1
    res[-1, ixNonRest] <- apply(tmpRet + 1, 2, cumprod) * 100
  } 
  # ixRate <- grepl("rate.", names(tmp), ignore.case=T)
  # if ( sum(ixRate) > 0 ) {
  #   res[1, ixRate] <- 100
  #   tmpRet <- tmp[-1, ixRate] - head(tmp[, ixRate], dim(tmp)[1]-1)
  #   res[-1, ixRate] <- apply(tmpRet, 2, cumsum) + 100
  # } 
  res
})
## traning sample DELTA according to the selected economic variables
df.train.sample.delta.VAR <- reactive({
  tmp <- df.full.sample.delta.VAR()
  filter(tmp, date > min(df.train.sample.VAR()[, 1]),
         date <= max(df.train.sample.VAR()[, 1]))
})
## training data for VAR(p) model fit
df.train.data2VAR.fit <- reactive({
  if (input$data2Fit.sel.VAR == "valOriginal"){
    res <- df.train.sample.VAR()
  } else if (input$data2Fit.sel.VAR == "valAjusted") {
    tmp <- df.train.sample.VAR()
    if (sum(input$delta.sel.VAR %in% c("asis", "diffAbs"))==1) {
        res <- tmp
    } else if (input$delta.sel.VAR == "diffLogAbs") {
        ixNonRest <- (!grepl("rate.", names(tmp), ignore.case=T)) & (names(tmp) != "date")
        res <- tmp
        res[, ixNonRest] <- log(res[, ixNonRest])
    } else if (input$delta.sel.VAR == "diffLog") {
        res <- tmp
        res[, names(tmp) != "date"] <- log(tmp[, names(tmp) != "date"])
    } else{
        stop("It is not defined here!!!")
    }
  } else if (input$data2Fit.sel.VAR == "valLogOrg"){
    res <- df.train.sample.VAR()
    res[, names(res) != "date"] <- log(res[, names(res) != "date"])
  } else {
    stop("Expand code for this case!!!")
  }
  res 
})
### US Recession stress period for training sample
df.UsRecess.train.VAR <- reactive({
  tmp <- df.train.data2VAR.fit()$date
  filter(dateUsRecess, monBeg >= min(tmp), monEnd <= max(tmp))
})
## testing sample according to the selected economic variables
df.test.sample.VAR <- reactive({
  tmp <- df.full.sample.VAR()
  Check.ExistVarInDF(tmp, "date")
  filter(tmp, date >= max(date.range.test()$dateBeg, min(tmp$date)),
         date <= min(date.range.test()$dateEnd, max(tmp$date)))
})
## testing sample DELTA according to the selected economic variables
df.test.sample.delta.VAR <- reactive({
  tmp <- df.full.sample.delta.VAR()
  filter(tmp, date > min(df.test.sample.VAR()[, 1]),
         date <= max(df.test.sample.VAR()[, 1]))
})
## testing data for VAR(p) model forecast
df.test.data2VAR.fit <- reactive({
  if (input$data2Fit.sel.VAR == "valOriginal"){
    res <- df.test.sample.VAR()
  } else if (input$data2Fit.sel.VAR == "valAjusted") {
    tmp <- df.test.sample.VAR()
    if (sum(input$delta.sel.VAR %in% c("asis", "diffAbs"))==1) {
      res <- tmp
    } else if (input$delta.sel.VAR == "diffLogAbs") {
      ixNonRest <- (!grepl("rate.", names(tmp), ignore.case=T)) & (names(tmp) != "date")
      res <- tmp
      res[, ixNonRest] <- log(res[, ixNonRest])
    } else if (input$delta.sel.VAR == "diffLog") {
      res <- tmp
      res[, names(tmp) != "date"] <- log(tmp[, names(tmp) != "date"])
    } else{
      stop("It is not defined here!!!")
    }
  } else if (input$data2Fit.sel.VAR == "valLogOrg"){
    res <- df.test.sample.VAR()
    res[, names(res) != "date"] <- log(res[, names(res) != "date"])
  } else {
    stop("Expand code for this case!!!")
  }
  res
})
df.test.sample.alignedVal.VAR <- reactive({
  tmp <- df.test.sample.VAR()
  res <- tmp
  ixNonRest <- (!grepl("rate.", names(tmp), ignore.case=T)) & (names(tmp) != "date")
  if (sum(ixNonRest) > 0){
    res[1, ixNonRest] <- 100
    tmpRet <- tmp[-1, ixNonRest] / head(tmp[, ixNonRest], dim(tmp)[1]-1) -1
    res[-1, ixNonRest] <- apply(tmpRet + 1, 2, cumprod) * 100
  } 
  res
})
### US Recession stress period for test sample
df.UsRecess.test.VAR <- reactive({
  tmp <- df.test.data2VAR.fit()$date
  filter(dateUsRecess, monBeg >= min(tmp), monEnd <= max(tmp))
})
################################################################END: reactive data prepare####

#################################BEGIN: reactive ANALYSES##################################
################################################################END: reactive ANALYSES#####


#################################BEGIN: Reactive model fitting #################################
### model fit for optimal lag
mod.VAR.algo <- reactive({
  VARselect(df.train.data2VAR.fit()[, names(df.train.data2VAR.fit()) != "date"], lag.max = 12, type = "both")
})
### VAR model fit
mod.VAR.select <- reactive({
  VAR(df.train.data2VAR.fit()[, names(df.train.data2VAR.fit()) != "date"], 
            p=as.numeric(input$sel.VAR.lag), type = "both")
})
### VAR model parameters

lst.modEst.VAR.select <- reactive({
  tmp <- summary(mod.VAR.select())
  nam <- names(tmp$varresult)
  tab.est <- data.frame(tmp$varresult[[1]]$coefficients[, 1], stringsAsFactors=F)
  tab.pVal <- data.frame(tmp$varresult[[1]]$coefficients[, 4], stringsAsFactors=F)
  vec.resid.sigma <- tmp$varresult[[1]]$sigma
  if(length(nam)>1){
    for (idx in 2:length(nam)){
      tmpDf <- data.frame(tmp$varresult[[idx]]$coefficients[, c(1, 4)], stringsAsFactors=F)
      tab.est <- cbind(tab.est, tmpDf[, 1])
      tab.pVal <- cbind(tab.pVal, tmpDf[, 2])
      vec.resid.sigma <- c(vec.resid.sigma, tmp$varresult[[idx]]$sigma)
    }
  }
  tab.est <- data.frame(tab.est, stringsAsFactors = F)
  tab.pVal <- data.frame(tab.pVal, stringsAsFactors = F)
  names(tab.est) <- nam
  names(tab.pVal) <- nam
  mat.cov <- tmp$covres
  mat.corr <- tmp$corres
  colnames(mat.cov) <- nam
  colnames(mat.corr) <- nam
  row.names(mat.cov) <- nam
  row.names(mat.corr) <- nam
  vec.eigen <- tmp$roots
  list(df.est=tab.est, df.est.pVal=tab.pVal, vec.sigma.stdErr=vec.resid.sigma, 
       vec.eigen=vec.eigen, mat.cov=mat.cov, mat.corr=mat.corr)
})


## VAR forecast for the input value
df.VAR.forecast.modOut <- reactive({
  tmp <- predict(mod.VAR.select(), n.ahead = dim(df.test.data2VAR.fit())[1])
  namVar <- names(tmp$fcst)
  df.pred <- data.frame(tmp$fcst[[1]][, 1:3], stringsAsFactors=FALSE)
  names(df.pred) <- paste0(namVar[1], c("_predict", "_LowerB", "_UpperB"))
  if (length(namVar) > 1) {
   for(idx in 2:length(namVar)){
     tst <- data.frame(tmp$fcst[[idx]][, 1:3], stringsAsFactors=FALSE)
     names(tst) <- paste0(namVar[idx], c("_predict", "_LowerB", "_UpperB"))
     df.pred <- cbind(df.pred, tst)
   }
  }
  df.pred
})
### combine the model forecast with the training sample (scale: VAR.fit, i.e., log value for delta of log diff)
df.VAR.forecast.val2VAR <- reactive({
  df.real <- df.test.data2VAR.fit()
  ### attached to the training sample for comparison
  df2Use <- df.test.data2VAR.fit()
  names(df2Use) <- c("date", paste0(names(df.real)[names(df.real) != "date"], "_value"))
  df2Use <- cbind(df2Use, df.VAR.forecast.modOut())
  df2Use
})
################################################################END: Reactive model fitting#####
```

Column {.tabset .tabset-fade .tabset-pills}
-----------------------------------------------------------------------
### Selected variables
#### Value of chosen variable
```{r}
DT::renderDataTable({
  df.train.sample.VAR()
  }, options = list(pageLength = 3, autoWidth = TRUE, rownames= FALSE)
)
```

#### $\Delta$ of chosen variable
```{r}
DT::renderDataTable({
  df.train.sample.delta.VAR()
  }, options = list(pageLength = 3, autoWidth = TRUE, rownames= FALSE)
)
```

#### Values used to fit VAR(p) model
```{r}
DT::renderDataTable({
  df.train.data2VAR.fit()
  }, options = list(pageLength = 3, autoWidth = TRUE, rownames= FALSE)
)
```


### Plot of time series 
- Plot of value over time 
```{r}
renderPlot({
  tmp <- df.train.data2VAR.fit()
  namDf <- names(tmp)[names(tmp) != "date"]
  eval(parse(text=paste0(
  "df2Plot <- tmp %>%",
    "gather(measure, value,", namDf[1], ":", tail(namDf, 1), ")"
  )))
  gObj <- ggplot(data=df2Plot) +
  geom_line(aes(x=date,y=value, group=measure), col="blue") +
  geom_point(aes(x=date,y=value, group=measure)) +
  facet_wrap(~measure, scales="free")
  if (nrow(df.UsRecess.train.VAR())==0) {
    gObj
  } else{
    gObj + geom_rect(data=df.UsRecess.train.VAR(),
            aes(xmin=monBeg, xmax=monEnd, ymin=-Inf, ymax=+Inf), fill='red', alpha=0.2)
  }
})
```

- Plot of $\Delta$ over time
```{r}
renderPlot({
  tmp <- df.train.sample.delta.VAR()
  namDf <- names(tmp)[names(tmp) != "date"]
  eval(parse(text=paste0(
  "df2Plot <- tmp %>%",
    "gather(measure, value,", namDf[1], ":", tail(namDf, 1), ")"
  )))
  gObj <- ggplot(data=df2Plot) +
  geom_line(aes(x=date,y=value, group=measure), col="blue") +
  geom_point(aes(x=date,y=value, group=measure)) +
  facet_wrap(~measure, scales="free")
  if (nrow(df.UsRecess.train.VAR())==0) {
    gObj
  } else{
    gObj + geom_rect(data=df.UsRecess.train.VAR(),
            aes(xmin=monBeg, xmax=monEnd, ymin=-Inf, ymax=+Inf), fill='red', alpha=0.2)
  }
})
```


### Stationary test
#### Stationarity test on the chosen variable and its **$\Delta$**

- [ADF](https://en.wikipedia.org/wiki/Augmented_Dickey%E2%80%93Fuller_test) test ($H_0$ of unit root presence, i.e.,  $I(1)$ process vs. $H_A$ stationary)

- [PP](https://en.wikipedia.org/wiki/Augmented_Dickey%E2%80%93Fuller_test) test ($H_0$ of unit root presence, i.e.,  $I(1)$ process v.s., $H_A$ stationary)

- [KPSS](https://en.wikipedia.org/wiki/KPSS_test) test ($H_0$ of stationary or $I(0)$ process v.s. $H_A$ non-stationary): 
This test can be performed assuming the deterministic component of the variable is a constant or a constant with a linear trend.

#### Test conclusions
```{r}
DT::renderDataTable({
  tmp <- df.train.sample.VAR()
  res.val <- TestsStationaryPval(tmp[, names(tmp) != "date"], k=2)
  tmp <- df.train.sample.delta.VAR()
  res.delta <- TestsStationaryPval(tmp[, names(tmp) != "date"], k=2)
  res.pVal <- cbind(res.val, res.delta)
  res.con <- apply(res.pVal, 2, function(x) ifelse(x < 0.05, "Stationary", "Non-stationary"))
  res.conKpss <- apply(res.pVal, 2, function(x) ifelse(x < 0.05, "Non-stationary", "Stationary"))
  res.con[, grepl("KPSS", colnames(res.con), ignore.case=T)] <- res.conKpss[, grepl("KPSS", colnames(res.con), ignore.case=T)]
  colnames(res.con) <- c("Value: ADF", "Value: PP", "Value: KPSS (trend)", "Value: KPSS (level)", 
                       "Delta: ADF", "Delta: PP", "Delta: KPSS (trend)", "Delta: KPSS (level)")
  row.names(res.con) <- names(tmp)[-1]
  res.con
  }, options = list(pageLength = 3, autoWidth = TRUE)
)  
```

#### Test p-values
``` {r}
DT::renderDataTable({
  tmp <- df.train.sample.VAR()
  res.val <- TestsStationaryPval(tmp[, names(tmp) != "date"], k=2)
  tmp <- df.train.sample.delta.VAR()
  res.delta <- TestsStationaryPval(tmp[, names(tmp) != "date"], k=2)
  res.pVal <- apply(cbind(res.val, res.delta), 2, function(x) round(x, 3))
  colnames(res.pVal) <- c("Value: ADF", "Value: PP", "Value: KPSS (trend)", "Value: KPSS (level)", 
                       "Delta: ADF", "Delta: PP", "Delta: KPSS (trend)", "Delta: KPSS (level)")
  row.names(res.pVal) <- names(tmp)[-1]
  res.pVal
  }, options = list(pageLength = 3, autoWidth = TRUE)
)  
```

#### Are the chosen time series $I(1)$?
If so, we can proceed to VAR fit 



### VAR Model
- Optimal lag of VAR(p) by statistical criteria
```{r}
renderPrint({
  tmp <- data.frame(t(as.numeric(mod.VAR.algo()$selection)), stringsAsFactors = F)
  names(tmp) <- names(mod.VAR.algo()$selection)
  tmp
})
```

- VAR(p) model fit given chosen p lag
```{r}
DT::renderDataTable({
  tmp <- lst.modEst.VAR.select()
  tab2Show <- paste0(round(tmp$df.est[, 1], 2), " (", round(tmp$df.est.pVal[, 1], 3),")")
  if (dim(tmp$df.est)[2] > 1){
    for (idx in 2:dim(tmp$df.est)[2]){
     tab2Show <- cbind(tab2Show, paste0(round(tmp$df.est[, idx], 2), " (", round(tmp$df.est.pVal[, idx], 3),")"))
    }
  }
  tab2Show <- data.frame(tab2Show, stringsAsFactors = F)
  names(tab2Show) <- paste0(names(tmp$df.est), ": Est.(pVal)")
  row.names(tab2Show) <- row.names(tmp$df.est)
  tab2Show
}, options = list(pageLength = 3, autoWidth = TRUE)
)
```

- Correlation matrix of model residuals
```{r}
DT::renderDataTable({
  tmp <- summary(mod.VAR.select())
  round(tmp$corres, 3)
}, options = list(pageLength = 2, autoWidth = TRUE)
)
```

- Causality test: Please note that the test is to determine whether one time series is useful in forecasting another.
```{r}
DT::renderDataTable({
  tmp <- mod.VAR.select()
  namVar <- colnames(tmp$y)
  res <- causality(tmp, cause=namVar[1])
  tab2Show <- data.frame(pVal.gran=round(as.numeric(res$Granger$p.value), 3), h0.gran=res$Granger$method, 
                         pVal.inst=round(as.numeric(res$Instant$p.value), 3), h0.inst=res$Instant$method, stringsAsFactors=FALSE)
  if (length(namVar) > 1){
    for (idx in 2:length(namVar)){
      res <- causality(tmp, cause=namVar[idx])
      tab2Show <- rbind(tab2Show, data.frame(pVal.gran=round(as.numeric(res$Granger$p.value), 3), h0.gran=res$Granger$method, 
                                             pVal.inst=round(as.numeric(res$Instant$p.value), 3), h0.inst=res$Instant$method, stringsAsFactors=FALSE))
    }
  }
  names(tab2Show) <- c("Granger p-value", "Granger Null hypothesis", "Instant p-value", "Instant Null hypothesis")
  tab2Show
  }, options = list(pageLength = 1, autoWidth = TRUE), rownames = FALSE
)
```


### Model Historical fit
```{r}
renderPlot({
  tmp <- mod.VAR.select()
  namVar <- colnames(tmp$y)
  df.Y <- data.frame(tmp$y, stringsAsFactors = F)
  df.Y$date <- df.train.data2VAR.fit()$date
  df.Y <- df.Y[, match(c("date", namVar), names(df.Y))]
  df.Y.fitted <- data.frame(tmp$varresult[[1]]$fitted.values, stringsAsFactors = F)
  if (length(namVar) > 1){
    for (idx in 2:length(namVar)){
      df.Y.fitted <- cbind(df.Y.fitted,
                           data.frame(tmp$varresult[[idx]]$fitted.values, stringsAsFactors = F))
    }
  }
  names(df.Y.fitted) <- namVar
  df.Y.fitted$date <- tail(df.Y$date, dim(df.Y.fitted)[1])
  df.Y.fitted <- df.Y.fitted[, match(c("date", namVar), names(df.Y.fitted))]
  eval(parse(text=paste0(
  "df2Plot.Y <- df.Y %>%",
    "gather(measure, value,", namVar[1], ":", tail(namVar, 1), ")"
  )))
  eval(parse(text=paste0(
  "df2Plot.Y.fitted <- df.Y.fitted %>%",
    "gather(measure, value,", namVar[1], ":", tail(namVar, 1), ")"
  )))
  gObj <- ggplot(data=df2Plot.Y) +
    geom_line(aes(x=date, y=value, group=measure), col="blue") +
    geom_point(aes(x=date, y=value, group=measure)) +
    facet_wrap(~measure, scales="free")+ 
    geom_line(data=df2Plot.Y.fitted, aes(x=date, y=value, group=measure)) +
    geom_line(data=df2Plot.Y.fitted, aes(x=date, y=value, group=measure), col="red") 
  if (nrow(df.UsRecess.train.VAR())==0) {
    gObj
  } else{
    gObj + geom_rect(data=df.UsRecess.train.VAR(),
            aes(xmin=monBeg, xmax=monEnd, ymin=-Inf, ymax=+Inf), fill='red', alpha=0.2)
  }
})
```


### Residual Tests

- Stationary test for individual residual series
```{r}
DT::renderDataTable({
  tmp <- mod.VAR.select()
  namVar <- colnames(tmp$y)
  df.mod.resid <- data.frame(tmp$varresult[[1]]$residuals, stringsAsFactors = F)
  if (length(namVar > 1)){
    for (idx in 2:length(namVar)){
      df.mod.resid <- cbind(df.mod.resid, 
                           data.frame(tmp$varresult[[idx]]$residuals, stringsAsFactors = F))
    }
  }
  names(df.mod.resid) <- namVar
  res.pVal <- TestsStationaryPval(df.mod.resid[, names(df.mod.resid) != "date"], k=2)
  res.con <- apply(res.pVal, 2, function(x) ifelse(x < 0.05, "Stationary", "Non-stationary"))
  res.conKpss <- apply(res.pVal, 2, function(x) ifelse(x < 0.05, "Non-stationary", "Stationary"))
  res.con[, grepl("KPSS", colnames(res.con), ignore.case=T)] <- res.conKpss[, grepl("KPSS", colnames(res.con), ignore.case=T)]
  colnames(res.con) <- c("Resid: ADF", "Resid: PP", "Resid: KPSS (trend)", "Resid: KPSS (level)")
  row.names(res.con) <- namVar
  res.con
  }, options = list(pageLength = 2, autoWidth = TRUE)
)  
```

- Residual **Autocorrelation**: 
The p-value of Portmanteau Test is `r renderUI(HTML(round(serial.test(mod.VAR.select(), type="PT.asymptotic")$serial$p.value, 3)))`, which implies that there is `r renderUI(HTML(ifelse(serial.test(mod.VAR.select(), type="PT.asymptotic")$serial$p.value < 0.05, " ", "NO")))` statistically significant autocorrelation among the model residuals. The p-value of Breusch-Godfrey LM Test is `r renderUI(HTML(round(serial.test(mod.VAR.select(), type="BG")$serial$p.value, 3)))`, which implies that there is `r renderUI(HTML(ifelse(vars::serial.test(mod.VAR.select(), type="BG")$serial$p.value < 0.05, " ", "NO")))` statistically significant autocorrelation among the model residuals.

- Residual **Heteroskedasticity**: The p-value of ARCH-LM test is `r renderUI(HTML(round(arch.test(mod.VAR.select())$arch.mul$p.value, 3)))`. Hence, there is `r renderUI(HTML(ifelse(arch.test(mod.VAR.select())$arch.mul$p.value < 0.05, " ", "NO ")))` statistically significant heteroskedasticity in the model residuals.


- Residuals over time
```{r}
renderPlot({
  tmp <- mod.VAR.select()
  namVar <- colnames(tmp$y)
  df.mod.resid <- data.frame(tmp$varresult[[1]]$residuals, stringsAsFactors = F)
  if (length(namVar > 1)){
    for (idx in 2:length(namVar)){
      df.mod.resid <- cbind(df.mod.resid, 
                           data.frame(tmp$varresult[[idx]]$residuals, stringsAsFactors = F))
    }
  }
  names(df.mod.resid) <- namVar
  df.mod.resid$date <- tail(df.train.data2VAR.fit()$date, dim(df.mod.resid)[1])
  eval(parse(text=paste0( 
    "df2Plot <- df.mod.resid %>%", 
    "gather(measure, value,", namVar[1], ":", tail(namVar, 1), ")" 
  )))
  gObj <- ggplot(data=df2Plot) +
    geom_point(aes(x=date, y=value, group=measure), col="blue", size=2) +
    facet_wrap(~measure, scales="free") + 
    geom_hline(yintercept = 0, col="red")
  if (nrow(df.UsRecess.train.VAR())==0) {
    gObj
  } else{
    gObj + geom_rect(data=df.UsRecess.train.VAR(),
            aes(xmin=monBeg, xmax=monEnd, ymin=-Inf, ymax=+Inf), fill='red', alpha=0.2)
  }
})
```

### Residuals ACF
```{r}
renderPlot({
  acf(residuals(mod.VAR.select()))
})
```

### Model Forecast
```{r}
renderPlot({
  tmp <- df.VAR.forecast.val2VAR()
  namDf <- names(tmp)[names(tmp) != "date"]
  eval(parse(text=paste0(
  "df2Plot <- tmp %>% ",
    "gather(measure, value,", namDf[1], ":", tail(namDf, 1), ")"
  )))
  pos.char <- unlist(regexec("_", as.character(df2Plot$measure)))
  df2Plot$variable <- NA
  df2Plot$type <- NA
  Check.StopIf(!identical(length(pos.char), dim(df2Plot)[1]), "Must be the same length here!!!")
  for (idx in 1:dim(df2Plot)[1]){
   df2Plot$variable[idx] <- substr(as.character(df2Plot$measure[idx]), 1, pos.char[idx]-1)
   df2Plot$type[idx] <- substr(as.character(df2Plot$measure[idx]), pos.char[idx]+1, nchar(as.character(df2Plot$measure[idx])))
  }
  ## DF contains model prediction and UB and LB of confi interval
  df2Ribbon <- df2Plot[, names(df2Plot) != "measure"] %>% # select(-measure) %>%  #!!! does not work
    filter(type != "value") %>% 
    spread(type, value)
  ## plot test sample
  gObj <- ggplot(data = filter(df2Plot, type == "value")) +
    geom_line(aes(x=date, y=value, group=variable), col="green") +
    geom_point(aes(x=date, y=value, group=variable), size=2, col="green") +
    facet_wrap(~variable, scales="free")
  ## overlay the forecasts and CI 
  gObj <- gObj + geom_line(data=df2Ribbon, aes(x=date, y=predict), col="blue") +
    geom_point(data=df2Ribbon, aes(x=date, y=predict), col="blue", size=2) + 
    geom_ribbon(data=df2Ribbon, aes(x=date, ymin = LowerB, ymax = UpperB), fill = "grey70", alpha=0.6)   #+
    #facet_wrap(~variable)
  ## overlay recession period if necessanary
  if (nrow(df.UsRecess.test.VAR())==0) {
    gObj
  } else{
    gObj + geom_rect(data=df.UsRecess.test.VAR(),
            aes(xmin=monBeg, xmax=monEnd, ymin=-Inf, ymax=+Inf), fill='red', alpha=0.2)
  }
})
```


### Cointegration 
[Johansen] test is performed to examine existence of conintegration among the chosen time series. Two tests (namely, Trace and Maximum Eigenvalue) are both performed, given the chosen lag of the fitted VAR(p) model.
``` {r}
# renderPrint({
#   #mat.inp <- matrix(, nrow=dim(df.train.data2VAR.fit())[1])
#   # tst <- attributes(summary(ca.jo(data.matrix(tmp), spec="transitory")))
#   # # tmp
#   # class(tst)
#   # tst
# })
```



ESG-VAR
=======================================================================

Column {.tabset .tabset-fade .tabset-pills}
-----------------------------------------------------------------------
### VAR(p) Parameter Override
```{r}
renderUI({
    tmp <- lst.modEst.VAR.select()
    vec.def.est <- round(tmp$df.est, 2)
    vec.def.name <- row.names(tmp$df.est.pVal)
    nCol <- dim(tmp$df.est)[2]
    txt2Eval <- purrr::map_chr(1:nCol, function(idx){
      paste0("renderNumInputs(vec.def.est[, ", idx, "], paste0(vec.def.name, '.coeff.', ", idx, "), vec.def.name, names(tmp$df.est[", idx, "]))")
    })
    txt2Eval <- paste(txt2Eval, collapse = ", ")
    eval(parse(text=paste0(
      "fillRow(", txt2Eval, ")"
    )))
})
```


```{r} 
df_sim_VAR <- eventReactive(input$buttSimVAR, {
  tmpEst <- lst.modEst.VAR.select()
  mCovMat <- as.matrix(tmpEst$mat.corr)
  varName.initial <- row.names(tmpEst$df.est.pVal)
  if (input$selOverrideVAR == "Original.Estiamte"){
    vCont <- as.numeric(tmpEst$df.est[varName.initial == "const", ])
    vTrend <- as.numeric(tmpEst$df.est[varName.initial == "trend", ])
    mEst <- t(as.matrix(tmpEst$df.est[!(varName.initial %in% c("const", "trend")), ]))
  } else if (input$selOverrideVAR == "Overriden"){
    nVarinY <- dim(lst.modEst.VAR.select()$df.est)[2]
    varNameConst <- paste0("const.coeff.", 1:nVarinY)
    varNameTrend <- paste0("trend.coeff.", 1:nVarinY)
    varNameCoeff <- unlist(purrr::map(1:(length(varName.initial)-2), 
                                      function(x) paste0(paste0(varName.initial[x], ".coeff."), 1:nVarinY)))
    ## get mEst
    txt2Eval <-paste("input$", varNameCoeff, collapse = ", ")
    eval(parse(text=paste0(
      paste0("mEst <- matrix(c(", txt2Eval, "), nrow=", nVarinY, ")")
    )))
    ## get vCont
    txt2Eval <-paste("input$", varNameConst, collapse = ", ")
    eval(parse(text=paste0(
      paste0("vCont <- as.numeric(c(", txt2Eval, "))")
    )))
    ## get vTrend
    txt2Eval <-paste("input$", varNameTrend, collapse = ", ")
    eval(parse(text=paste0(
      paste0("vTrend <- as.numeric(c(", txt2Eval, "))")
    )))
  } else{
    stop("Only can take into account Original.Estiamte or Overriden.")
  }
  ### check whether the VAR coefficient make senses
  Check.InputCoeff.VAR(mEst)
  chol(mCovMat) ### check whether covariance matrix is psd
  ### trend matrix to add
  nPeriod2Fore <- input$modHorizonByInputPeriod + 1
  mTrend <- matrix(1:nPeriod2Fore, ncol = 1) %*% matrix(vTrend, nrow=1)
  res <- matrix(NA, nrow = input$nrSimVAR, ncol = nPeriod2Fore*dim(mEst)[1])
  for (idx in 1:input$nrSimVAR){
    res[idx, ] <- unlist(c(Sim_VAR_N_Period(w=vCont, A=mEst, C=mCovMat, N=nPeriod2Fore) + mTrend ))
  }
  res <- data.frame(res, stringsAsFactors = F)
  names(res) <- unlist(purrr::map(names(tmpEst$df.est), function(x) paste0(x, "_Period", 1:nPeriod2Fore)))
  res
})

df_sim_VAR_Forecast_Input<- reactive({
  ### delta should be perform over columns 
  uniVarName <- names(df.full.sample.VAR())[!(toupperNoSpace(names(df.full.sample.VAR())) %in% c("DATE", "DATES"))]
  Check.Unique(uniVarName)
  ### get initial value input
  txt2Ev <- paste(paste0("input$", uniVarName, "_foreT0"), collapse = ", ")
  eval(parse(text=paste0(
    "vecInitial <- c(", txt2Ev, ")"
  )))
  names(vecInitial) <- uniVarName
  ### prepare value to input in 
  if (input$data2Fit.sel.VAR == "valOriginal" | (input$data2Fit.sel.VAR == "valAjusted" & input$delta.sel.VAR %in% c("asis", "diffAbs"))) {
    vecInput <- vecInitial
  } else if (input$data2Fit.sel.VAR == "valLogOrg" | (input$data2Fit.sel.VAR == "valAjusted" & input$delta.sel.VAR == "diffLog")){
    vecInput <- log(vecInitial)
  } else if (input$data2Fit.sel.VAR == "valAjusted" & input$delta.sel.VAR == "diffLogAbs"){
    ixNonRest <- (!grepl("rate.", names(vecInitial), ignore.case=T)) & (names(vecInitial) != "date")
    vecInput <- vecInitial
    vecInput[ixNonRest] <- log(vecInput[ixNonRest])
  } else {
    stop("Expand code for this case!!!")
  }
  ### calculate cumulative sum delta for each variable across differnet forecast period
  nPeriodAhead <- input$modHorizonByInputPeriod + 1
  res <- data.frame(matrix(NA, nrow = dim(df_sim_VAR())[1], ncol = length(uniVarName)*(nPeriodAhead-1)), stringsAsFactors = F)
  names(res) <- unlist(purrr::map(uniVarName, function(x) paste0(x, "_Period", 1:(nPeriodAhead-1))))
  for (idx in 1:length(uniVarName)){
    Check.ExistVarInDF(df_sim_VAR(), paste0(uniVarName[idx], "_Period", 1:nPeriodAhead))
    res[, (1 + (idx-1)*(nPeriodAhead-1)):((nPeriodAhead-1)*idx)] <- t(apply(
      (df_sim_VAR()[, match(paste0(uniVarName[idx], "_Period", 2:nPeriodAhead), names(df_sim_VAR()))] -
      df_sim_VAR()[, match(paste0(uniVarName[idx], "_Period", 1:(nPeriodAhead-1)), names(df_sim_VAR()))]) , 1, cumsum)) +
    vecInput[names(vecInput) == uniVarName[idx]]
  }
  # ### show forecasted values
  if (input$data2Fit.sel.VAR == "valOriginal" | (input$data2Fit.sel.VAR == "valAjusted" & input$delta.sel.VAR %in% c("asis", "diffAbs"))) {
    res
  } else if (input$data2Fit.sel.VAR == "valLogOrg" | (input$data2Fit.sel.VAR == "valAjusted" & input$delta.sel.VAR == "diffLog")){
    exp(res)
  } else if (input$data2Fit.sel.VAR == "valAjusted" & input$delta.sel.VAR == "diffLogAbs"){
    ixNonRest <- (!grepl("rate.", names(res), ignore.case=T)) & (names(res) != "date")
    if (any(ixNonRest)){
      res[, ixNonRest] <- exp(res[, ixNonRest])
    }
    res
  } else {
    stop("Expand code for this case!!!")
  }
  
})
```


<!-- ### debug simValue -->
<!-- ```{r} -->
<!-- renderPrint({ -->
<!--   print(df_sim_VAR()) -->
<!-- }) -->
<!-- ``` -->


### debug forecast
```{r}
renderPrint({
  print(df_sim_VAR_Forecast_Input())
})
```



### VAR(p) Simulation
``` {r}
nVarinY <- 3
fillRow(width="100%", flex = c(NA, 1), 
  wellPanel(
    fluidRow(
      selectInput("selOverrideVAR", label="VAR parameters to use", choices = c("Original.Estiamte", "Overriden")),
      sliderInput(inputId = "modHorizonByInputPeriod", label = "Nr of periods ahead to Sim", min = 2, max = 36, value = 20),
      sliderInput(inputId = "nrSimVAR", label = "Nr of simulation", min = 100, max = 10e3, value = 500),
      sliderInput(inputId = "nrBinHistVAR", label = "Nr of bins for histogram", min = 1, max = 30, value = 10),
      sliderInput(inputId = "period2ShowHist", label = "n-th period to show in histgram", min = 1, max = 36, value = 6), 
      actionButton("buttSimVAR", "Simulate VAR")
    )
  ),
  plotOutput("barTest", height = "100%"),
  fluidRow(  
    column(12, renderUI({
      tmp <- tail(df.full.sample.VAR()[, -1], 1)
      renderNumInputs(as.numeric(tmp), paste0(names(tmp), "_foreT0"), names(tmp), "Initial value for forecast")
      })
    )
  )
)

### make a test plot
output$barTest <- renderPlot({
  Check.StopIf(input$period2ShowHist > input$modHorizonByInputPeriod, paste0("We can maximum show input$modHorizonByInputPeriod","-th period!" ))
  tmp <- df_sim_VAR_Forecast_Input()
  tmp[, grepl(paste0('_Period', input$period2ShowHist), names(tmp))] %>% 
    gather(Variable, Value) %>% 
    ggplot(aes(x=Value)) +
    geom_histogram(aes(y = ..density..), alpha=0.9, col="blue", bins=input$nrBinHistVAR) +
    geom_density(col="red", size=1.1, alpha=0.6) +
    facet_wrap(~Variable, scales="free")
  #plotly(gObj_bar)
})
```


### VAR sim verification

```{r}
### Verification for lag of 1
renderPrint({
  summary(VAR(df_sim_VAR()[1, ], p=input$sel.VAR.lag, type="both"))
})
```
